---
title: "Albacore MSE"
subtitle: "MP v OM"
author: "Laurie"
date: "05/04/2015"
output: html_document
---

```{r, echo=FALSE}
library(knitr)

opts_chunk$set(comment=NA, fig.width =6, 
                           fig.height=4,
                           fig.path  ="../tex/",
                           warning=FALSE, 
                           message=FALSE, 
                           error  =FALSE, 
                           echo   =FALSE,
                           cache  =!FALSE)
```

```{r}
library(FLCore)
library(ggplotFL)
library(reshape)
library(plyr)
library(kobe)
library(FLBRP)
library(biodyn)
library(popbio)

theme_set(theme_bw(base_size=10))

dirDat="c:/temp/albn/data"
dirDat="/home/laurie/MEGAsync/mse/albn/data"

load(paste(dirDat,"/om.RData",sep=""))

r<-function(x){
  f   =c(FLBRP:::refpts(x)["crash","harvest"])
  if (is.na(f)) 
      f=c(refpts(x)["msy","harvest"])*3
                                  
  r=try(log(lambda(leslie(x,f))))
  if (class(r)=="try-error") 
      r=NA  
  r}

priorFn<-function(x,p=NULL){
  
  # production function based ob S/R
  msy =c(FLBRP:::refpts(x)["msy",   "yield"])
  bmsy=c(FLBRP:::refpts(x)["msy",   "biomass"])
  
  r=r(x)
  k   =c(FLBRP:::computeRefpts(x)["virgin","biomass"])
  
  if (is.null(p))
  p   =optimise(function(p,bmsy,k) 
                (bmsy-k*(1/(1+p))^(1/p))^2, c(0.001,5), 
                bmsy=bmsy,k=k)$minimum

  kPrime=bmsy/((1/(1+p))^(1/p))
  rPrime=msy/(k*(1/(1+p))^(1/p+1))
  
  cbind(p=p,
        r     =r,     k     =k,
        rPrime=rPrime,kPrime=kPrime,
        msy=msy,bmsy=bmsy,fmsy=msy/bmsy)}
```

```{r}
FLBRP2biodyn=function(from){
  
  ## age based regference points
  msy =c(from@refpts["msy",   "yield"])
  bmsy=c(from@refpts["msy",   "biomass"])
  k   =c(from@refpts["virgin","biomass"])
  
  # biomass based reference points
  p   =optimise(function(p,bmsy,k) 
    (bmsy-k*(1/(1+p))^(1/p))^2, c(0.001,5), bmsy=bmsy,k=k)$minimum
  k=bmsy/((1/(1+p))^(1/p))
  r=msy/(k*(1/(1+p))^(1/p+1))
  b0=mean(stock(from)[,1:5]/k)
  
  bd=biodyn()
  
  bd@params=FLPar(c(r=r,k=k,p=p,b0=b0))
  
  bd@catch=catch.obs(from)
  bd@stock=from@stock.obs
  
  range(bd)["minyear"]=dims(bd@catch)$minyear
  range(bd)["maxyear"]=dims(bd@catch)$maxyear
  
  bd}

fitBd<-function(om,eql,priors,
                fixQ=FALSE,prr="",
                rMult=1,b0Mult=1,kMult=1,
                rtn=function(x,y){
                 model.frame(mcf(FLQuants(
                            omF=catch(x)/stock(x),
                            omB=stock(x),
                            mpF=harvest(y),
                            mpB=stock(y))),drop=T)})
  {
  bd=FLBRP2biodyn(eql)
  catch(bd)=catch(om)
  
  params(bd)[c("r","k","p","b0")]=unlist(c(priors[,c("rPrime","k","p","b0")]))
  params(bd)["r" ]=params(bd)["r" ]*rMult
  params(bd)["k" ]=params(bd)["k" ]*kMult
  params(bd)["b0"]=params(bd)["b0"]*b0Mult
  
  bd@stock=FLQuant(c(params(bd)["k"]*params(bd)["b0"]),
                   dimnames=list(age="all",year=dimnames(stock(om))$year))
  bd=window(bd,end=dims(stock(om))$maxyear-1)
  
  bd=fwd(bd,catch=catch(bd))
  
  u =window((stock(om)[,- dim(stock(om))[2]]+
             stock(om)[,-1])/2,start=start)

  setParams(bd)=u
  setControl(bd)=params(bd)
  
  if (fixQ)
    control(bd)["q1",c("phase","val")]=c(-1,1)

  if (prr=="q"){
    bd@priors["q1","weight"]=1
    bd@priors["q1","a"]     =1 
  }else if (prr=="bmsy"){
    bd@priors["bmsy","weight"]=1
    bd@priors["bmsy","a"]     =priors[,"bmsy"]
  }else if (prr=="fmsy"){
    bd@priors["fmsy","weight"]=1
    bd@priors["fmsy","a"]     =priors[,"fmsy"]
  }else if (prr=="msy"){
    bd@priors["msy","weight"]=1
    bd@priors["msy","a"]     =priors[,"msy"]
  }else if (prr=="r"){
    bd@priors["r","weight"]=1
    bd@priors["r","a"]     =priors[,"r"]
  }else if (prr=="k"){
    bd@priors["k","weight"]=1
    bd@priors["k","a"]     =priors[,"k"]
  }else if (prr=="rPrime"){
    bd@priors["r","weight"]=1
    bd@priors["r","a"]     =priors[,"rPrime"]
  }else if (prr=="kPrime"){
    bd@priors["k","weight"]=1
    bd@priors["k","a"]     =priors[,"kPrime"]}

  bd=fit(bd,u)

  res=rtn(om,bd)
  
#   ggplot(res)+
#     geom_line(aes(year,omB,group=i))+
#     geom_line(aes(year,mpB,group=i),col="red")

  res
  }

goodFn<-function(det){
  
  rms=ddply(det,.(i,rMult,fixQ,prr),
            function(x) sum(((x$mpB-x$omB)/x$omB)^2,na.rm=T)/length(x$omB))
  
  rms =rms[do.call(order,rms[,c("i","V1")]),]
  good=rms[!duplicated(rms[,"i"]),]
  
  good[do.call(order,list(V1=good[,"V1"])),]}
```

```{r-albn-sa-det}
start=1975

priors=ldply(eql,priorFn,p=NULL)
priors=cbind(priors,mdply(seq(length(om)), 
              function(i) 
                data.frame(b0=mean(stock(window(om[[i]],end=start))[,1:10]))))[,-10]
priors=transform(priors,b0=pmin(b0/k,0.8))

runs=expand.grid(i    =seq(length(om)),
                 rMult=1.5,
                 fixQ =!as.logical(TRUE),
                 prr  ="")

det=mdply(runs,
          function(i,rMult,fixQ,prr){
            #print(paste(i,rMult,prr))
                fitBd(window(om[[i]],start=start),
                      eql[[i]],priors[i,],fixQ=TRUE,rMult=rMult)})
ggplot(det)+
  geom_line(aes(year,omB,group=i))+
  geom_line(aes(year,mpB,group=i),col="red")
```

**Figure 1** Deterministic comparison between OM and MP for p known and truncated time series.

```{r-albn-sa-stochastic,fig.height=8}
sa=biodyns(mlply(runs,
          function(i,rMult,fixQ,prr){
            #print(paste(i,rMult,prr))
                fitBd(window(om[[i]],start=start),
                      eql[[i]],priors[i,],fixQ=TRUE,rMult=rMult,
                      rtn=function(x,y) y)}))

u=FLQuants(llply(om,function(x) window(stock(x),start=start)))
u=FLQuants(llply(u, function(x) rlnorm(100,log(x),.2)))

sa1975=biodyns(mlply(seq(length(sa)), function(i) {
  catch(sa[[i]])=propagate(catch(sa[[i]]),100)
  fit(sa[[i]],u[[i]])}))
names(sa1975)=names(om)

save(sa1975,u,file=paste(dirDat,"sa1975.RData",sep="/"))
```

```{r}
ggplot(rbind.fill(cbind(What="MP",plot(FLQuants(llply(sa,stock)),na.rm=TRUE)$data),
                  cbind(What="OM",plot(FLQuants(llply(om,stock)),na.rm=TRUE)$data)))+
       geom_ribbon(aes(year,min=`10%`,max=`90%`,fill=What),alpha=.5)+
       geom_line(aes(year,`50%`,col=What))+
       facet_wrap(~qname,ncol=2)

```

**Figure 2a** Comparison of stochastic CPUE MP and deterministic OM for p known and truncated time series.


```{r-albn-sa-p1-det}
ggplot(rbind.fill(
       cbind(What="MP",
            plot(FLQuants(mlply(names(sa),function(x) 
              stock(sa[[x]])%/%bmsy(sa[[x]]))))$data),
       cbind(What="OM",
             plot(FLQuants(mlply(names(om), function(x) stock(om[[x]])%/%FLBRP:::refpts(eql[[x]])["msy","biomass"])))$data)))+
       geom_ribbon(aes(year,min=`10%`,max=`90%`,fill=What),alpha=.5)+
       geom_line( aes(year,`50%`,col=What))+
       geom_hline(aes(yintercept=1))+
       facet_wrap(~qname,ncol=2)
```

**Figure 2b** Comparison of stochastic CPUE MP and deterministic OM for p known and truncated time series.


```{r-albn-sa-p1-det}
start=1975

priors=ldply(eql,priorFn,p=1)
priors=cbind(priors,mdply(seq(length(om)), 
              function(i) 
                data.frame(b0=mean(stock(window(om[[i]],end=start))[,1:10]))))[,-10]
priors=transform(priors,b0=pmin(b0/k,0.8))

runs=expand.grid(i    =seq(length(om)),
                 rMult=1.5,
                 fixQ =!as.logical(TRUE),
                 prr  ="")

det=mdply(runs,
          function(i,rMult,fixQ,prr){
            #print(paste(i,rMult,prr))
                fitBd(window(om[[i]],start=start),
                      eql[[i]],priors[i,],fixQ=TRUE,rMult=rMult)})
ggplot(det)+
  geom_line(aes(year,omB,group=i))+
  geom_line(aes(year,mpB,group=i),col="red")
```

**Figure 3** Deterministic comparison between OM and MP for p known and truncated time series.

```{r-albn-sa-p1-stochastic,fig.height=8}
sa=biodyns(mlply(runs,
          function(i,rMult,fixQ,prr){
            #print(paste(i,rMult,prr))
                fitBd(window(om[[i]],start=start),
                      eql[[i]],priors[i,],fixQ=TRUE,rMult=rMult,
                      rtn=function(x,y) y)}))

u=FLQuants(llply(om,function(x) window(stock(x),start=start)))
u=FLQuants(llply(u, function(x) rlnorm(100,log(x),.2)))

sa1975p1=biodyns(mlply(seq(length(sa)), function(i) {
  catch(sa[[i]])=propagate(catch(sa[[i]]),100)
  fit(sa[[i]],u[[i]])}))
names(sa1975p1)=names(om)

save(sa1975p1,u,file=paste(dirDat,"sa1975.RData",sep="/"))

ggplot(rbind.fill(cbind(What="MP",plot(FLQuants(llply(sa1975p1,stock)),na.rm=TRUE)$data),
                  cbind(What="OM",plot(FLQuants(llply(om,      stock)),na.rm=TRUE)$data)))+
       geom_ribbon(aes(year,min=`10%`,max=`90%`,fill=What),alpha=.5)+
       geom_line(aes(year,`50%`,col=What))+
       facet_wrap(~qname,ncol=2)
```

**Figure 4** Comparison of stochastic CPUE MP and deterministic OM for p known and truncated time series.


```{r-albn-sa1950-p1-det}
start=1950

priors=ldply(eql,priorFn,p=NULL)
priors=cbind(priors,mdply(seq(length(om)), 
              function(i) 
                data.frame(b0=mean(stock(window(om[[i]],start=start))[,1:10]))))[,-10]
priors=transform(priors,b0=pmin(b0/k,0.8))

runs=expand.grid(i    =seq(length(om)),
                 rMult=1.5,
                 fixQ =!as.logical(TRUE),
                 prr  ="")

det=mdply(runs,
          function(i,rMult,fixQ,prr){
            #print(paste(i,rMult,prr))
                fitBd(window(om[[i]],start=start),
                      eql[[i]],priors[i,],fixQ=TRUE,rMult=rMult)})
ggplot(det)+
  geom_line(aes(year,omB,group=i))+
  geom_line(aes(year,mpB,group=i),col="red")
```

**Figure 5** Deterministic comparison between OM and MP for p known and truncated time series.

```{r-albn-sa1950-p1-stochastic,fig.height=8}
sa=biodyns(mlply(runs,
          function(i,rMult,fixQ,prr){
            #print(paste(i,rMult,prr))
                fitBd(window(om[[i]],start=start),
                      eql[[i]],priors[i,],fixQ=TRUE,rMult=rMult,
                      rtn=function(x,y) y)}))

u=FLQuants(llply(om,function(x) window(stock(x),start=start)))
u=FLQuants(llply(u, function(x) rlnorm(100,log(x),.2)))

sa1950=biodyns(mlply(seq(length(sa)), function(i) {
  catch(sa[[i]])=propagate(catch(sa[[i]]),100)
  fit(sa[[i]],u[[i]])}))
names(sa1950)=names(om)

save(sa1950,u,file=paste(dirDat,"sa1950.RData",sep="/"))

ggplot(rbind.fill(cbind(What="MP",plot(FLQuants(llply(sa1950,stock)))$data),
                  cbind(What="OM",plot(FLQuants(llply(om,    stock)))$data)))+
       geom_ribbon(aes(year,min=`10%`,max=`90%`,fill=What),alpha=.5)+
       geom_line(aes(year,`50%`,col=What))+
       facet_wrap(~qname,ncol=2)
```

**Figure 6** Comparison of stochastic CPUE MP and deterministic OM for p known and truncated time series.


```{r-albn-sa1950-det}
start=1950

priors=ldply(eql,priorFn,p=1)
priors=cbind(priors,mdply(seq(length(om)), 
              function(i) 
                data.frame(b0=mean(stock(window(om[[i]],start=start))[,1:10]))))[,-10]
priors=transform(priors,b0=pmin(b0/k,0.8))

runs=expand.grid(i    =seq(length(om)),
                 rMult=1.5,
                 fixQ =!as.logical(TRUE),
                 prr  ="")

det=mdply(runs,
          function(i,rMult,fixQ,prr){
            #print(paste(i,rMult,prr))
                fitBd(window(om[[i]],start=start),
                      eql[[i]],priors[i,],fixQ=TRUE,rMult=rMult)})
ggplot(det)+
  geom_line(aes(year,omB,group=i))+
  geom_line(aes(year,mpB,group=i),col="red")
```

**Figure 7** Deterministic comparison between OM and MP for p known and truncated time series.

```{r-albn-sa1950-stochastic,fig.height=8}
sa=biodyns(mlply(runs,
          function(i,rMult,fixQ,prr){
            #print(paste(i,rMult,prr))
                fitBd(window(om[[i]],start=start),
                      eql[[i]],priors[i,],fixQ=TRUE,rMult=rMult,
                      rtn=function(x,y) y)}))

u=FLQuants(llply(om,function(x) window(stock(x),start=start)))
u=FLQuants(llply(u, function(x) rlnorm(100,log(x),.2)))

sa1950p1=biodyns(mlply(seq(length(sa)), function(i) {
  catch(sa[[i]])=propagate(catch(sa[[i]]),100)
  fit(sa[[i]],u[[i]])}))
names(sa1950p1)=names(om)

save(sa1950p1,u,file=paste(dirDat,"sa1950.RData",sep="/"))

ggplot(rbind.fill(cbind(What="MP",plot(FLQuants(llply(sa1950p1,stock)))$data),
                  cbind(What="OM",plot(FLQuants(llply(om,      stock)))$data)))+
       geom_ribbon(aes(year,min=`10%`,max=`90%`,fill=What),alpha=.5)+
       geom_line(aes(year,`50%`,col=What))+
       facet_wrap(~qname,ncol=2)
```

**Figure 8** Comparison of stochastic CPUE MP and deterministic OM for p known and truncated time series.


```{r-albn-sa-prj,fig.height=8,eval=FALSE}
srDev=rlnorm(100,FLQuant(0,dimnames=list(year=2012:2041)))
o=fwdWindow(om[["Base"]],end=2041,eql[["Base"]])
o=fwd(o,f=FLQuant(c(FLBRP:::refpts(eql[["Base"]])["msy","harvest"]*0.7),
                            dimnames=list(year=2012:2041)),
                            sr=eql[["Base"]],
                            sr.residuals=srDev*.6)

names(counterfactual)=names(om)          
plot(FLQuants(llply(counterfactual[-6],stock))) +
  facet_wrap(~qname,scale="free")

u=FLQuants(llply(om,function(x) window(stock(x),start=start)))
u=FLQuants(llply(u, function(x) rlnorm(100,log(x),.2)))

load(paste(dirDat,"sa.RData",sep="/"))
sa=biodyns(mlply(seq(length(sa)), function(i) {
  catch(sa[[i]])=propagate(catch(sa[[i]]),100)
  fit(sa[[i]],u[[i]])}))
names(sa)=names(om)

ggplot(rbind.fill(cbind(What="MP",plot(FLQuants(llply(sa,stock)))$data),
                  cbind(What="OM",plot(FLQuants(llply(om,stock)))$data)))+
       geom_ribbon(aes(year,min=`10%`,max=`90%`,fill=What),alpha=.5)+
       geom_line(aes(year,`50%`,col=What))+
       facet_wrap(~qname,ncol=2)
```

**Figure 9** Comparison of stochastic CPUE MP and deterministic OM with regime shift.
