---
title: "North Atlanic Albacore"
subtitle: "Implicit Management Procedure"
author: "Laurence Kell"
date: "26/03/2015"
output: html_document
---

```{r, echo=FALSE}
library(knitr)

opts_chunk$set(comment=NA, 
               fig.width =6, 
               fig.height=6,
               fig.path  ="../tex/",
               warning=FALSE, 
               message=FALSE, 
               error  =FALSE, 
               echo   =FALSE, 
               eval   =TRUE,
               cache  =!TRUE)
```

```{r, echo=FALSE}
#install.packages("C:/Users/laurie.kell/Documents/MEGAsync/temp/aspic_2.0.1.tar.gz",repos=NULL,type="source",INSTALL_opts=c("--no-multiarch"))

library(FLCore)
library(ggplotFL)
library(scales)
library(plyr)
library(reshape)
library(FLBRP)
library(biodyn)
library(aspic)
library(plotrix)
library(corrplot)
library(diags)

theme_set(theme_bw(10))

dirMy ="c:/temp/albn"
dirDat=paste(dirMy,"/data",sep="")
dirInp="http://rscloud.iccat.int/kobe/Inputs/albn/2013/aspic/"
```

```{r}
# ASPIC runs
#time series 
ts=mdply(paste(dirInp,"/run",1:7,"/aspic.rdat",sep=""),
          function(x) data.frame(biomass=dget(x)$t.series$b,
                                 harvest=dget(x)$t.series$F.total,
                                 year   =dget(x)$t.series$year))
```

```{r}
# Parameter estimates
pnms=c("power","B1.K","MSY","Fmsy","Bmsy","K","r","phi","b.bmsy","F.Fmsy","yield.eq")

par=mdply(paste(dirInp,"/run",1:7,"/aspic.rdat",sep=""),
          function(x) dget(x)$estimates[pnms])
```

```{r}
# aspic objects

asp=aspics(paste(dirInp,"/run",1:7,"/aspic.inp",sep=""))
control(asp[[4]])["k","val"]=par[4,"K"]
control(asp[[4]])["msy","val"]=par[4,"MSY"]
control(asp[[4]])["k","fit"]=0

asp=aspics(llply(asp,fit))
```

```{r}
## CPUE indices

idx=ldply(asp,index)
idn=ddply(idx,.(.id), with, data.frame(name=unique(name)))
idn=ddply(idn,.(.id),with, 
          data.frame(index=name,
                     name =paste("cpue0",seq(length(name)),sep="")))

names(idn)[1]="X1"

idx=idx[do.call(order,idx[,c("name","year")]),]
idx=idx[!duplicated(idx[,c("name","year")]),-c(1,4,6)]
idx=idx[!is.na(idx$index),]

## Diagnostics
dgs=mdply(paste(dirInp,"/run",1:7,"/aspic.prn",sep=""), function(x) diags(x,"aspic"))
dgs=merge(dgs,idn)
```

```{rabn-imp-cpue, fig.width=8, fig.height=12}
library(gam)

gm  =gam(log(obs)~lo(year)+index,data=dgs)
dgs=data.frame(dgs,gam=predict(gm),gamdgsl=residuals(gm))
scl =coefficients(gm)[-(1:2)]
#names(scl)=substr(names(scl),5,nchar(names(scl)))
dgs=transform(dgs,scl=scl[as.character(name)])
dgs[is.na(dgs$scl),"scl"]=0

ggplot(dgs)+ geom_line(aes(year,exp(gam)),col="red") +
             geom_smooth(aes(year,obs),se=FALSE)      +           
             geom_point(aes(year,obs,col=index))       +
             theme(legend.position="bottom")          +
             theme(legend.position="none")            +
             facet_wrap(~index,ncol=1,scale="free_y")  +
             theme_ms(legend.position="none")
```

**Figure 1** CPUE indices, blue line is a lowess smoother by series, red line a GAM fitted to all data.

```{rabn-imp-cpueCor}
cr=cor(cast(idx,year~name,value="index")[,-1],use="pairwise.complete.obs")

nms=names(cast(idx,year~name,value="index")[,-1])
dimnames(cr)=list(nms,nms)
cr[is.na(cr)]=0
corrplot(cr,diag=F,order="hclust",addrect=2)  +          
             theme(legend.position="bottom")  
```

**Figure 2** Correlation between indices


```{rabn-imp-tsB,fig.height=4,fig.width=8}
ggplot(ts)+
  geom_line(aes(year,biomass,group=X1,col=X1))+
  kobe:::theme_ms(legend.position="none")
```

**Figure 3** Time series of biomass from 2013 assessment.


```{rabn-imp-tsF,fig.height=4,fig.width=8}
ggplot(ts)+
  geom_line(aes(year,harvest,group=X1,col=X1))+
  kobe:::theme_ms(legend.position="none")
```

**Figure 4** Time series of biomass from 2013 assessment.


```{rabn-imp-aspic,fig.height=4,fig.width=6}
plot(asp)+theme_bw()
```

**Figure 5** ASPIC runs


```{rabn-imp-diags}
ggplot(ddply(dgs, .(index,X1), with, data.frame(obs=stdz(obs),hat=stdz(hat))))+
          geom_abline(aes(0,1))                         +
          geom_point( aes(obs,hat,group=index,col=index))                     +
          stat_smooth(aes(obs,hat),method="lm", se=F)    +
          #facet_grid(~method,scale="free")                                   +
          theme_ms(14,legend.position="none")            +
          xlab("Fitted") + ylab("Observed")+
          facet_grid(X1~index)
```

**Figure 6** CPUE fits


```{rabn-imp-cf,eval=FALSE,fig.height=4,fig.width=8}
i=c(1,2,4,5)[1] #Troll, CT, JLL, All
# 1&5 OK, 2&4 bad

bd  =as(asp[[i]],"biodyn")
cpue=index(asp[[i]],FALSE)
control(bd)[c("r"),"val"]=1.2*params(bd)[c("r")]
bd1=fwd(bd,catch=catch(bd))
bd1=fit(bd1,cpue)

plot(biodyns("ASPIC"=bd,"biodyn"=bd1))+
  theme_bw()
```


```{rabn-imp-aspicOps,fig.height=4,fig.width=6}
lag<-function(x) {
  res=FLQuant(NA,dimnames=dimnames(x))
  res[,-dim(res)[2]][]=c(x[,-1])
  
  res}
i=2
tm=model.frame(mcf(FLQuants(
         lag    =lag(stock(asp[[i]])),
         biomass=stock(asp[[i]]),
         catch  =catch(asp[[i]]))),drop=T)
tm=transform(tm,prd=lag-biomass+catch)
ggplot(tm)+
  geom_path(aes(biomass,prd,col=year))+
  geom_point(aes(biomass,prd,col=year))+
  theme_bw()
```

**Figure 7** Productivity for run 2

```{rabn-imp-xval}
xval=function(object,cpue,minyear,maxyear=range(object)["maxyear"]){

  u =window(cpue,    end=minyear)
  bd=window(object,  end=minyear)
  bd=fit(bd,u)
  bd=fwd(bd,catch=catch(object)[,ac((minyear+1):maxyear)])
  
  nU     =(dim(params(bd))[1]-4)/2
  biomass=(stock(bd)[,-dim(stock(bd))[2]]+stock(bd)[,-1])/2
  
  if (!("FLQuants"%in%is(cpue))) cpue=FLQuants(cpue)
  
  res=mdply(seq(nU), function(i)
   model.frame(mcf(FLQuants(
          hat=biomass%*%params(bd)[4+i],
          obs=cpue[[i]])),drop=T))

  subset(res,year%in%minyear:maxyear&!is.na(obs))}
```

## Cross-validation

```{rabn-imp-xval1,fig.height=4,fig.height=6}
i=5
bd  =as(asp[[i]],"biodyn")
cpue=index(asp[[i]],FALSE)
bd1=fwd(bd,catch=catch(bd))
bd1=fit(bd1,cpue)

x=xval(bd,cpue,2000)

ggplot(x)+geom_point(aes(obs,hat))+
  facet_wrap(~X1,ncol=2,scale="free")+
  theme_bw()
```

**Figure 8** Correlation between observed and predicted CPUE.

```{rabn-imp-xval2}
ggplot(x)+geom_point(aes(year,obs))+
          geom_line( aes(year,hat))+
  facet_wrap(~X1,ncol=1,scale="free")+
  theme_bw()
```

**Figure 9** Comparison between observed and predicted CPUEs.


```{rabn-imp-xval3}
with(subset(x,X1==unique(X1)[2]),taylor.diagram(log(obs),log(hat),pos.cor=FALSE,normalize=TRUE,col="cyan",ref.sd=TRUE))

d_ply(x,.(X1),with, taylor.diagram(log(obs),log(hat),col="black",
                                            add=TRUE,pos.cor=TRUE,
                                            normalize=TRUE))
```

**Figure 10** Taylor Diagrammes

